---
title: "Introduction to hotteR"
author: "Trent Henderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to hotteR}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 7,
  fig.width = 9
)
```

```{r, message = FALSE, warning = FALSE}
library(hotteR)
```

`hotteR` is an R package that enables automated analysis of the [Triple J Hottest 100](https://www.abc.net.au/triplej/hottest100/20/) and [triplejplays](https://twitter.com/triplejplays) data in R.

## Core functions

### get_countdowns()

You can webscrape all prior Hottest 100 Countdown results (including Hottest 100 of the decade) using `get_countdowns()`. Note that you must have Chrome version 87.0.4280.88 and Java Development Kit installed. This function is somewhat unstable on Mac and may error out at a random year due to some known issues with `RSelenium`.

```{r message = FALSE, eval = FALSE, warning = FALSE}
countdown_data <- get_countdowns()
```

### historical_countdowns

A full dataset of all previous Hottest 100 Countdowns available to-date comes ready with the package. It has columns for rank, year, artist name, song name, and artist nationality.

```{r message = FALSE, warning = FALSE}
head(historical_countdowns)
```

### do_countdown_analysis()

You can automatically get some high level analysis and data visualisations for data retrieved from `get_countdowns()` or `historical_countdowns` using `do_countdown_analysis()`.

```{r message = FALSE, warning = FALSE, fig.keep = TRUE}
do_countdown_analysis(historical_countdowns)
```

### get_plays()

You can automatically pull summed counts of plays from the @triplejplays Twitter account using the `get_plays()` one-line function. The data is cleaned, processed, and summed behind the scenes, leaving you more time to focus on analysis and data visualisation.

*NOTE: This function uses access to the Twitter API and thus requires appropriate credentials to be available in the R Environment prior to running. The ['twitteR'](https://github.com/geoffjentry/twitteR) package provides a useful setup_twitter_oauth() function for this.*

```{r message = FALSE, eval = FALSE, warning = FALSE}
play_data <- get_plays(year = 2020)
```

### do_plays_analysis()

You can automatically get some high level analysis and data visualisations for data retrieved from `get_plays()` using `do_plays_analysis()`.

```{r message = FALSE, eval = FALSE, warning = FALSE}
play_data <- get_plays(year = 2020)
do_plays_analysis(the_plays)
```

### plot_probabilities()

You can automatically produce density plots of probability of membership for each Hottest 100 Countdown quartile and nationality with `plot_probabilities()`. This function takes one of two timescales as an argument:

* "Last Decade"
* "All Time"

An example for "Last Decade" is shown below.

```{r message = FALSE, warning = FALSE, fig.keep = TRUE}
plot_probabilities(timescale = "Last Decade")
```

### plot_quartile_area()

You can automatically produce time series area plots by quartile and nationality using `plot_quartile_area()`:

```{r message = FALSE, warning = FALSE, fig.keep = TRUE}
plot_quartile_area()
```

### theme_hotteR()

As well as data scraping and analysis functions, `hotteR` also provides a distinct and striking contrasting colour palette that is called and used by default when the package is loaded.

```{r message = FALSE}
scales::show_col(hotteR:::hotteR_palette)
```
